Effectful Semantics for Substructural SSA

### 2023-11-22 pldi24/paper.tex
Work on terms with holes semantics
Tap...

\begin{theorem}[Weakening] \
  \begin{enumerate}
    \item If \(\cwk{\Gamma}{\Delta}\) and \(q \subseteq r\),
    \(\hasty{\Delta}{p}{a}{A}{r}\), then \(\hasty{\Gamma}{p}{a}{A}{q}\)
    \item If \(\cwk{\Gamma}{\Delta}\), \(\haslb{\Delta}{t}{\ms{L}}\), then
    \(\haslb{\Gamma}{t}{\ms{L}}\)
    \item If \(\lwk{\ms{L}}{\ms{K}}\), \(\haslb{\Gamma}{t}{\ms{L}}\) then
    \(\haslb{\Gamma}{t}{\ms{K}}\)
  \end{enumerate}
\end{theorem}

\begin{proof}
  By induction: 
  \begin{enumerate}
    \item
    \item
    \item 
  \end{enumerate}
\end{proof}

\begin{lemma}[Substitution Splitting] 
  If \(\csplits{\Gamma}{\Delta}{\Xi}\) and \(\issubst{\gamma}{\Theta}{\Gamma}\),
  then there exist \(\Theta_\Delta\), \(\Theta_\Xi\) such that:
  \begin{itemize}
    \item \(\csplits{\Theta}{\Theta_\Delta}{\Theta_\Xi}\)
    \item \(\issubst{\gamma}{\Theta_\Delta}{\Delta}\)
    \item \(\issubst{\gamma}{\Theta_\Xi}{\Xi}\)
  \end{itemize}
\end{lemma}

@@

\begin{lemma}[Substitution] 
  Given \(\issubst{\gamma}{\Theta}{\Gamma}\),
  \begin{itemize}
    \item If \(\hasty{\Gamma}{p}{a}{A}{q}\), then \(\hasty{\Theta}{p}{[\gamma]a}{A}{q}\)
    \item If \(\haslb{\Gamma}{t}{\ms{L}}\), then there exists \(\ms{K}\) such that \(\lbsubst{\substlbs{\gamma}{\ms{L}}}{\ms{K}}{\ms{L}}\) and \(\haslb{\Theta}{[\gamma]t}{\ms{K}}\)
    \item If \(\lhaslb{\ms{W}}{L}{\ms{L}}\), then there exists \(\ms{K}, \ms{W}'\) such that 
    \begin{itemize}
      \item \(\lbsubst{\substlbs{\gamma}{\ms{L}}}{\ms{K}}{\ms{L}}\) 
      \item \(\lbsubst{\substlbs{\gamma}{\ms{W}'}}{\ms{W}}{\ms{W}'}\) 
      \item \(\lhaslb{\ms{W}'}{[\gamma]L}{\ms{K}}\)
    \end{itemize}
  \end{itemize}
\end{lemma}

@@

\begin{lemma}[Semantic Upgrade]
  For all \(\Gamma, a, A\), if \(\hasty{\Gamma}{1}{a}{A}{q}\), then
  \[\dnt{\hasty{\Gamma}{0}{a}{A}{q}} = \upg{\dnt{\hasty{\Gamma}{1}{a}{A}{q}}}{}\]
\end{lemma}

\begin{lemma}[Semantic Weakening] 
  If \(\cwk{\Gamma}{\Delta}\), then 
  \begin{itemize}
    \item For all \(\hasty{\Delta}{p}{a}{A}{q}\),
    \[
      \upg{\dnt{\cwk{\Gamma}{\Delta}}}{p}
      ; \dnt{\hasty{\Delta}{p}{a}{A}{q}}
      = \dnt{\hasty{\Gamma}{p}{a}{A}{q}} 
    \]
    \item For all \(\haslb{\Delta}{t}{\ms{L}}\),
    \[
      \upg{\dnt{\cwk{\Gamma}{\Delta}}}{p}
      ; \dnt{\haslb{\Delta}{t}{\ms{L}}}
      = \dnt{\haslb{\Gamma}{t}{\ms{L}}}  
    \]
  \end{itemize}

  Similarly, if \(\lwk{\ms{L}}{\ms{K}}\) and \(\haslb{\Gamma}{t}{\ms{L}}\), then
  \[
    \dnt{\haslb{\Gamma}{t}{\ms{L}}}
    ; \dnt{\lwk{\ms{L}}{\ms{K}}}
    = \dnt{\haslb{\Gamma}{t}{\ms{K}}}
  \]
\end{lemma}

@@

    <\(H\)> ::= \(\cdot\) 
    \;|\; \(H, \tyhole{\lhole{X}}{\Gamma}{p}{A}{q}\)
    \;|\; \(H, \blkhole{\lhole{T}}{\Gamma}{\ms{L}}\)
    \;|\; \(H, \cfghole{\lhole{W}}{\ms{L}}{\ms{K}}\)
  \end{grammar}
  \caption{Grammar for \isotopessa expressions and blocks with holes}
  \Description{Grammar for isotope-SSA blocks with holes}
  \label{fig:blocks-with-holes-grammar}
\end{figure}
We introduce a typing judgement \(\isrw{\mc{H}}{I}{H}\), which says that
\(\mc{H}\) assigns to each hole in \(H\) a term satisfying the corresponding
typing judgement within the hole context \(I\). We also introduce new typing
judgements for terms, blocks, and CFGs with holes in Figure
\ref{fig:ssa-holes-judgements}. In each case, they correspond to the judgement
that, given a collection \(\mc{H}\) terms, blocks, and CFGs to fill each hole in
\(H\), would satisfy the corresponding judgement without holes in
\ref{fig:ssa-judgements}.

@@

\begin{figure}
  \begin{gather*}    
    \prftree[r]{\rle{hole-tm}}
      {\tyhole{\lhole{X}}{\Delta}{p}{A}{q}}
      {\cwk{\Gamma}{\Delta}}
      {\mhasty{H}{\Gamma}{p}{\lhole{X}}{A}{q}} \qquad
    \prftree[r]{\rle{var}}
      {\cwk{\Gamma}{\thyp{x}{A}{q}}}
      {\mhasty{H}{\Gamma}{p}{x}{A}{q}} \qquad
    \prftree[r]{\rle{app}}
      {f \in \mc{I}_p^q(A, B)}
      {\mhasty{H}{\Gamma}{1}{\mhole{a}}{A}{q}}
      {\mhasty{H}{\Gamma}{p}{f\;\mhole{a}}{B}{q}} \\
    \prftree[r]{\rle{pair}}
      {\csplits{\Gamma}{\Delta}{\Xi}}
      {\mhasty{H}{\Delta}{1}{\mhole{a}}{A}{q}}
      {\mhasty{H}{\Xi}{1}{\mhole{b}}{B}{q}}
      {\mhasty{H}{\Gamma}{p}{(\mhole{a}, \mhole{b})}{A \otimes B}{q}} \\
    \prftree[r]{\rle{unit}}
      {\cwk{\Gamma}{\cdot}}
      {\mhasty{H}{\Gamma}{p}{()}{\mb{1}}{q}} \qquad
    \prftree[r]{\rle{true}}
      {\cwk{\Gamma}{\cdot}}
      {\mhasty{H}{\Gamma}{p}{\ctt}{\mb{2}}{q}} \qquad
    \prftree[r]{\rle{false}}
      {\cwk{\Gamma}{\cdot}}
      {\mhasty{H}{\Gamma}{p}{\cff}{\mb{2}}{q}} \\
    \prftree[r]{\rle{let}}
      {\csplits{\Gamma}{\Delta}{\Xi}}
      {\mhasty{H}{\Delta, \thyp{x}{A}{}}{p}{\mhole{e}}{B}{q}}
      {\mhasty{H}{\Xi}{1}{\mhole{a}}{A}{q}}
      {\mhasty{H}{\Gamma}{p}{\letexpr{x}{\mhole{a}}{\mhole{e}}}{B}{q}} \qquad
    % \prftree[r]{\rle{blk}}
    %   {\mhaslb{H}{\Gamma}{\mhole{t}}{\llhyp{\ell}{\cdot}{A}}}
    %   {\mhasty{H}{\Gamma}{0}{\lbsplice{\ell}{A}{\mhole{t}}}{A}{\varnothing}} 
      \\
    \prftree[r]{\rle{let2}}
      {\csplits{\Gamma}{\Delta}{\Xi}}
      {\mhasty{H}{\Delta, \thyp{x}{A}{}, \thyp{y}{B}{}}{p}{\mhole{e}}{C}{q}}
      {\mhasty{H}{\Xi}{1}{\mhole{a}}{A \otimes B}{q}}
      {\mhasty{H}{\Gamma}{p}{\letexpr{(x, y)}{\mhole{a}}{\mhole{e}}}{C}{q}}
  \end{gather*}
  \caption{Typing rules for \isotopessa terms with holes}
  \Description{Typing rules for isotope-SSA terms with holes}
  \label{fig:ssa-term-holes-typing}
\end{figure}

@@

We can interpret a hole context semantically as the \textit{set} of all possible
tuples of morphisms which would typecheck in place of the holes in the context,
as specified in Figure \ref{fig:holes-semantics}.
\begin{figure}
  \begin{align*}
      \\
    \dnt{H, \tyhole{\lhole{X}}{\Gamma}{p}{A}{q}}
      &= \dnt{H} \times \mc{C}_p^q(\dnt{\Gamma}, \dnt{A}) 
      \\
    \dnt{H, \blkhole{\lhole{T}}{\Gamma}{\ms{L}}}
      &= \dnt{H} \times \mc{C}_0(\dnt{\Gamma}, \dnt{\ms{L}}) 
      \\
    \dnt{H, \cfghole{\lhole{W}}{\ms{L}}{\ms{K}}}
      &= \dnt{H} \times \mc{C}_0(\dnt{\ms{L}}, \dnt{\ms{K}} + \dnt{\ms{L}}) 
      \\
    \dnt{\cdot} 
      &= \mb{1}
  \end{align*}
  \caption{Semantics for \isotopessa holes}
  \Description{Semantics for isotope-SSA holes}
  \label{fig:holes-semantics}
\end{figure}
We can now quite easily give semantics for terms, blocks, and CFGs with holes
which simply correspond to "filling in" any holes with the appropriate morphism
from the hole context, with all other rules interpreted identically to their
regular variants: this is done in Figures \ref{fig:ssa-term-holes-semantics} and
\ref{fig:ssa-block-holes-semantics}.

\begin{figure}
  \begin{gather*}
    \boxed{
      \dnt{\mhasty{H}{\Gamma}{p}{\mhole{a}}{A}{q}}
      : \dnt{H} \to \mc{C}_p^q(\dnt{\Gamma}, \dnt{A})}
      \\
    \sorry
  \end{gather*}
  \caption{Semantics for \isotopessa terms with holes}
  \Description{Semantics for isotope-SSA terms with holes}
  \label{fig:ssa-term-holes-semantics}
\end{figure}

>>>>

\begin{lemma} We have that
  \begin{enumerate}
    \item Splitting commutes: if \(\csplits{\Gamma}{\Delta}{\Xi}\) then \(\csplits{\Gamma}{\Xi}{\Delta}\), and \(\dnt{\csplits{\Gamma}{\Delta}{\Xi}} = \dnt{\csplits{\Gamma}{\Xi}{\Delta}};\sigma\)
    \item Splitting associates: if \(\csplits{\Gamma}{\Delta}{\Xi}\) and \(\csplits{\Xi}{\Theta}{\Phi}\) then there exists \(\Gamma_{\Delta, \Theta}\) such that \(\csplits{\Gamma}{\Gamma_{\Delta, \Theta}}{\Phi}\) and \(\csplits{\Gamma_{\Delta, \Theta}}{\Delta}{\Theta}\), and, for \emph{all} such \(\Gamma_{\Delta, \Theta}\), we have \(\dnt{\csplits{\Gamma}{\Delta}{\Xi}};\dnt{\Delta} \otimes \csplits{\Xi}{\Theta}{\Phi} = \dnt{\csplits{\Gamma}{\Gamma_{\Delta, \Theta}}{\Phi}};\dnt{\csplits{\Gamma_{\Delta, \Theta}}{\Delta}{\Theta}} \otimes \dnt{\Phi};\alpha\)
    \item Weakening composes: if \(\cwk{\Gamma}{\Delta}\) and \(\cwk{\Delta}{\Xi}\), then \(\cwk{\Gamma}{\Xi}\)
    \item Splitting and weakening compose on the right: if \(\csplits{\Gamma}{\Delta}{\Xi}\) and \(\cwk{\Delta}{\Delta'}\) then \(\csplits{\Gamma}{\Delta'}{\Xi}\), furthermore,
    \(\dnt{\csplits{\Gamma}{\Delta'}{\Xi}} = \dnt{\csplits{\Gamma}{\Delta}{\Xi}};\dnt{\cwk{\Delta}{\Delta'}} \otimes \dnt{\Xi}\)
    \item Splitting and weakening compose on the left: if \(\csplits{\Gamma}{\Delta}{\Xi}\) and \(\cwk{\Gamma'}{\Gamma}\) then \(\csplits{\Gamma'}{\Delta'}{\Xi}\), furthermore,
    \(\dnt{\csplits{\Gamma'}{\Delta}{\Xi}} = \dnt{\cwk{\Gamma'}{\Gamma}};\dnt{\csplits{\Gamma}{\Delta}{\Xi}}\)
  \end{enumerate}
\end{lemma}

\begin{proof}
  By induction.
\end{proof}

\begin{lemma}[Weakening] \
  \begin{enumerate}
    \item If \(\cwk{\Gamma}{\Delta}\) and \(q \subseteq r\),
    \(\hasty{\Delta}{p}{a}{A}{r}\), then \(\hasty{\Gamma}{p}{a}{A}{q}\)
    \item If \(\cwk{\Gamma}{\Delta}\), \(\haslb{\Delta}{t}{\ms{L}}\), then
    \(\haslb{\Gamma}{t}{\ms{L}}\)
    \item If \(\lwk{\ms{L}}{\ms{K}}\), \(\haslb{\Gamma}{t}{\ms{L}}\) then
    \(\haslb{\Gamma}{t}{\ms{K}}\)
  \end{enumerate}
\end{lemma}

\begin{proof}
  By induction.
\end{proof}

\begin{lemma}[Substitution Splitting] 
  If \(\csplits{\Gamma}{\Delta}{\Xi}\) and \(\issubst{\gamma}{\Theta}{\Gamma}\),
  then there exist \(\Theta_\Delta\), \(\Theta_\Xi\) such that
  \(\csplits{\Theta}{\Theta_\Delta}{\Theta_\Xi}\),
  \(\issubst{\gamma}{\Theta_\Delta}{\Delta}\), and
  \(\issubst{\gamma}{\Theta_\Xi}{\Xi}\). In particular, if
  \(\cwk{\Gamma}{\Delta}\) and \(\issubst{\gamma}{\Theta}{\Gamma}\), then
  \(\issubst{\gamma}{\Theta}{\Delta}\).
\end{lemma}

@@

\begin{lemma}[Substitution] 
  Given \(\issubst{\gamma}{\Theta}{\Gamma}\),
  \begin{enumerate}
    \item If \(\hasty{\Gamma}{p}{a}{A}{q}\), then \(\hasty{\Theta}{p}{[\gamma]a}{A}{q}\) \label{lemma:substitution:term}
    \item If \(\haslb{\Gamma}{t}{\ms{L}}\), then there exists \(\ms{K}\) such that \(\lbsubst{\substlbs{\gamma}{\ms{L}}}{\ms{K}}{\ms{L}}\) and \(\haslb{\Theta}{[\gamma]t}{\ms{K}}\)
    \label{lemma:substitution:block}
    \item If \(\lhaslb{\ms{W}}{L}{\ms{L}}\), then there exists \(\ms{K},
    \ms{W}'\) such that \(\lbsubst{\substlbs{\gamma}{\ms{L}}}{\ms{K}}{\ms{L}}\),
    \(\lbsubst{\substlbs{\gamma}{\ms{W}'}}{\ms{W}}{\ms{W}'}\), and
    \(\lhaslb{\ms{W}'}{[\gamma]L}{\ms{K}}\)
    \label{lemma:substitution:cfg}
  \end{enumerate}
\end{lemma}

\begin{proof} 
  We begin by proving \ref{lemma:substitution:term} by induction on derivations \(\hasty{\Gamma}{p}{a}{A}{q}\):
  \begin{itemize}
    \item \rle{var}: we have that \(\cwk{\Gamma}{\thyp{x}{A}{q}}\), and
    therefore that \(\issubst{\gamma}{\Theta}{\thyp{x}{A}{q}}\); by
    \rle{subst-var}, we obtain the desired result that
    \(\hasty{\Theta}{p}{\gamma(x)}{A}{q}\).
    \item \rle{app}: \todo{this}
    \item \rle{pair}: \todo{this}
    \item \rle{true}: \todo{this}
    \item \rle{false}: \todo{this}
    \item \rle{let}: \todo{this}
    \item \rle{unit}: \todo{this}
    \item \rle{let2}: \todo{this}
  \end{itemize}
  Using \ref{lemma:substitution:term}; we can now prove
  \ref{lemma:substitution:block} and \ref{lemma:substitution:cfg} by
  mutual induction on derivations as follows:
  \begin{itemize}
    \item \rle{br}: \todo{this}
    \item \rle{let-blk}: \todo{this}
    \item \rle{ite}: \todo{this}
    \item \rle{let2-blk}: \todo{this}
    \item \rle{where}: \todo{this}
    \item \rle{nil-br}: \todo{this}
    \item \rle{cons-br}: \todo{this}
  \end{itemize}
\end{proof}

\begin{proof} 
  We proceed by mutual induction on derivations as follows:
  \begin{itemize}
    \item \rle{br}: \todo{this}
    \item \rle{let-blk}: \todo{this}
    \item \rle{ite}: \todo{this}
    \item \rle{let2-blk}: \todo{this}
    \item \rle{where}: \todo{this}
    \item \rle{nil-br}: \todo{this}
    \item \rle{cons-br}: \todo{this}
  \end{itemize}
\end{proof}

\begin{lemma}[Semantic Upgrade]
  If \(\hasty{\Gamma}{1}{a}{A}{q}\), we have that
  \(\dnt{\hasty{\Gamma}{0}{a}{A}{q}} =
  \upg{\dnt{\hasty{\Gamma}{1}{a}{A}{q}}}{}\)
\end{lemma}

@@

\begin{lemma}[Semantic Weakening] 
  If \(\cwk{\Gamma}{\Delta}\), then 
  \begin{itemize}
    \item For all \(\hasty{\Delta}{p}{a}{A}{q}\),
    \(
      \upg{\dnt{\cwk{\Gamma}{\Delta}}}{p}
      ; \dnt{\hasty{\Delta}{p}{a}{A}{q}}
      = \dnt{\hasty{\Gamma}{p}{a}{A}{q}} 
    \)
    \item For all \(\haslb{\Delta}{t}{\ms{L}}\),
    \(
      \upg{\dnt{\cwk{\Gamma}{\Delta}}}{p}
      ; \dnt{\haslb{\Delta}{t}{\ms{L}}}
      = \dnt{\haslb{\Gamma}{t}{\ms{L}}}  
    \)
  \end{itemize}

  Similarly, if \(\lwk{\ms{L}}{\ms{K}}\) and \(\haslb{\Gamma}{t}{\ms{L}}\), then
  \(
    \dnt{\haslb{\Gamma}{t}{\ms{L}}}
    ; \dnt{\lwk{\ms{L}}{\ms{K}}}
    = \dnt{\haslb{\Gamma}{t}{\ms{K}}}
  \)
\end{lemma}

@@

    <\(H\)> ::= \(\cdot\) 
    \;|\; \(M\)

    <\(M\)> ::= \(\tyhole{\lhole{X}}{\Gamma}{p}{A}{q}\)
    \;|\; \(\blkhole{\lhole{T}}{\Gamma}{\ms{L}}\)
    \;|\; \(\cfghole{\lhole{W}}{\ms{L}}{\ms{K}}\)
  \end{grammar}
  \caption{Grammar for \isotopessa expressions and blocks with holes}
  \Description{Grammar for isotope-SSA blocks with holes}
  \label{fig:blocks-with-holes-grammar}
\end{figure}
We introduce a typing judgement \(\isrw{\mc{H}}{I}{H}\), which says that
\(\mc{H}\) assigns to each hole in \(H\) a term satisfying the corresponding
typing judgement within the hole context \(I\). We also introduce new typing
judgements for terms, blocks, and CFGs with holes in Figure
\ref{fig:ssa-holes-judgements}. In each case, they correspond to the judgement
that, given a collection \(\mc{H}\) terms, blocks, and CFGs to fill each hole in
\(H\), would satisfy the corresponding judgement without holes in
\ref{fig:ssa-judgements}.

@@

\begin{figure}
  \begin{gather*}    
    \prftree[r]{\rle{hole-tm}}
      {\tyhole{\lhole{X}}{\Delta}{p}{A}{q} \in H}
      {\cwk{\Gamma}{\Delta}}
      {\mhasty{H}{\Gamma}{p}{\lhole{X}}{A}{q}} \qquad
    \prftree[r]{\rle{var}}
      {\cwk{\Gamma}{\thyp{x}{A}{q}}}
      {\mhasty{H}{\Gamma}{p}{x}{A}{q}} \qquad
    \prftree[r]{\rle{app}}
      {f \in \mc{I}_p^q(A, B)}
      {\mhasty{H}{\Gamma}{1}{\mhole{a}}{A}{q}}
      {\mhasty{H}{\Gamma}{p}{f\;\mhole{a}}{B}{q}} \\
    \prftree[r]{\rle{pair}}
      {\csplits{\Gamma}{\Delta}{\Xi}}
      {\mhasty{H}{\Delta}{1}{\mhole{a}}{A}{q}}
      {\mhasty{H}{\Xi}{1}{\mhole{b}}{B}{q}}
      {\mhasty{H}{\Gamma}{p}{(\mhole{a}, \mhole{b})}{A \otimes B}{q}} \\
    \prftree[r]{\rle{unit}}
      {\cwk{\Gamma}{\cdot}}
      {\mhasty{H}{\Gamma}{p}{()}{\mb{1}}{q}} \qquad
    \prftree[r]{\rle{true}}
      {\cwk{\Gamma}{\cdot}}
      {\mhasty{H}{\Gamma}{p}{\ctt}{\mb{2}}{q}} \qquad
    \prftree[r]{\rle{false}}
      {\cwk{\Gamma}{\cdot}}
      {\mhasty{H}{\Gamma}{p}{\cff}{\mb{2}}{q}} \\
    \prftree[r]{\rle{let}}
      {\csplits{\Gamma}{\Delta}{\Xi}}
      {\mhasty{H}{\Delta, \thyp{x}{A}{}}{p}{\mhole{e}}{B}{q}}
      {\mhasty{H}{\Xi}{1}{\mhole{a}}{A}{q}}
      {\mhasty{H}{\Gamma}{p}{\letexpr{x}{\mhole{a}}{\mhole{e}}}{B}{q}} \qquad
    % \prftree[r]{\rle{blk}}
    %   {\mhaslb{H}{\Gamma}{\mhole{t}}{\llhyp{\ell}{\cdot}{A}}}
    %   {\mhasty{H}{\Gamma}{0}{\lbsplice{\ell}{A}{\mhole{t}}}{A}{\varnothing}} 
      \\
    \prftree[r]{\rle{let2}}
      {\csplits{\Gamma}{\Delta}{\Xi}}
      {\mhasty{H}{\Delta, \thyp{x}{A}{}, \thyp{y}{B}{}}{p}{\mhole{e}}{C}{q}}
      {\mhasty{H}{\Xi}{1}{\mhole{a}}{A \otimes B}{q}}
      {\mhasty{H}{\Gamma}{p}{\letexpr{(x, y)}{\mhole{a}}{\mhole{e}}}{C}{q}}
  \end{gather*}
  \caption{Typing rules for \isotopessa terms with holes}
  \Description{Typing rules for isotope-SSA terms with holes}
  \label{fig:ssa-term-holes-typing}
\end{figure}

@@

We can interpret a hole context semantically as the \textit{set} of all possible
tuples of morphisms which would typecheck in place of the holes in the context,
as specified in Figure \ref{fig:holes-semantics}.
\begin{figure}
  \begin{gather*}
    \dnt{\tyhole{\lhole{X}}{\Gamma}{p}{A}{q}} 
      = \mc{C}_p^q(\dnt{\Gamma}, \dnt{A})
    \qquad
    \dnt{\blkhole{\lhole{T}}{\Gamma}{\ms{L}}} 
      = \mc{C}_0(\dnt{\Gamma}, \dnt{\ms{L}})
      \\
    \dnt{\cfghole{\lhole{W}}{\ms{L}}{\ms{K}}}
      = \mc{C}_0(\dnt{\ms{L}}, \dnt{\ms{K}} + \dnt{\ms{L}}) 
      \\
    \dnt{H, M} = \dnt{H} \times \dnt{M}
    \qquad
    \dnt{\cdot} = \mb{1}
      \\
    \boxed{\dnt{M \in H}: \dnt{H} \to \dnt{M}}
      \qquad
    \dnt{M \in H, M \in H} 
      = \pi_1
    \qquad
    \dnt{M \in H, N} = \pi_0;\dnt{M \in H}
  \end{gather*}
  \caption{Semantics for \isotopessa holes}
  \Description{Semantics for isotope-SSA holes}
  \label{fig:holes-semantics}
\end{figure}
We can now quite easily give semantics for terms, blocks, and CFGs with holes
which simply correspond to "filling in" any holes with the appropriate morphism
from the hole context, with all other rules interpreted identically to their
regular variants: this is done in Figures \ref{fig:ssa-term-holes-semantics} and
\ref{fig:ssa-block-holes-semantics}.

\begin{figure}
  \begin{gather*}
    \boxed{
      \dnt{\mhasty{H}{\Gamma}{p}{\mhole{a}}{A}{q}}
      : \dnt{H} \to \mc{C}_p^q(\dnt{\Gamma}, \dnt{A})}
      \\
    \sorry
      \\
    \boxed{
      \dnt{\tyhole{\mhole{X}}{\Gamma}{a}{A}{q} \in H}
      : H \to \mc{C}^q_p(\dnt{\Gamma}, \dnt{A})
    }
  \end{gather*}
  \caption{Semantics for \isotopessa terms with holes}
  \Description{Semantics for isotope-SSA terms with holes}
  \label{fig:ssa-term-holes-semantics}
\end{figure}
